Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)

### 2025-08-06 resubmission/02-antecedents.qmd
Up antecedentes meritocracia + render
@juancarloscastillo todo tuyo

Meritocracy constitutes a central ideological framework for legitimizing different types of social inequality, for instance through market justice beliefs. Rooted in the belief that rewards and positions should be allocated based on individual effort and talent, meritocracy operates as a normative ideal and a descriptive belief about how society functions. As initially conceptualized by Michael Young [-@young_rise_1962], the term was meant to critique a system in which merit-based stratification becomes a new form of inequality. However, over time, meritocracy has been widely supported in many societies as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented economies [@mijs_paradox_2021; @sandel_tyranny_2020]. From a sociological perspective, the belief in meritocracy is more than a cognitive assessment; it reflects a moral lens through which individuals interpret inequality. People who believe that success results from hard work and talent are more likely to view social and economic disparities as legitimate [@castillo_percepcion_2019; @batruch_belief_2023]. Conversely, if they see outcomes as driven by luck, social origin, or systemic barriers, inequality is more likely to be perceived as unjust. This distinction becomes crucial in societies with persistent structural inequality, where public narratives often emphasize personal responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing between two key dimensions: effort-based and talent-based perceptions. This distinction is essential, as it captures different pathways through which individuals justify inequality. Effort-based meritocracy emphasizes hard work and perseverance as the basis for social rewards, aligning closely with cultural narratives of personal responsibility. A talent-based meritocracy, by contrast, emphasizes intelligence and innate abilities, which are often perceived as less malleable and more unequally distributed. Both dimensions have been shown to correlate with acceptance of inequality, but they may carry distinct implications for how inequality is justified in specific domains [@castillo_multidimensional_2023]. The relevance of this distinction is supported by recent studies, which show that individuals respond differently to these dimensions. For instance, perceptions that effort is rewarded in society are more strongly associated with positive evaluations of fairness and acceptance of unequal outcomes [@batruch_belief_2023].

These dimensions of meritocracy reflect how respondents perceive society's distributive logic, regardless of whether they endorse meritocratic principles. This distinction aligns with recent findings indicating that individuals distinguish between how merit is perceived in society and how it should ideally operate, which in turn shapes their preferences for redistribution and justice [@tejero-peregrina_perceived_2025]. Meritocratic beliefs serve as symbolic justifications for unequal outcomes, particularly when access is stratified by income or social opportunity. Prior studies have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes [@castillo_percepcion_2019; @castillo_socialization_2024].

In addition to influencing individual attitudes toward inequality, meritocratic perceptions can contribute to social division and the stigmatization of disadvantaged groups. Recent research has demonstrated that exposure to meritocratic narratives can reinforce the belief that poverty results from individual failings rather than systemic conditions, reducing support for redistributive measures and increasing the stigmatization of the poor [@hoyt_mindsets_2023]. This reinforces negative stereotypes and reduces empathy toward individuals from lower socioeconomic backgrounds. Moreover, @busemeyer_positive_2021 argues that meritocratic narratives can serve as feedback mechanisms that shape public opinion and well-being by framing individuals' understanding of welfare outcomes as deserved or undeserved within existing institutional structures. This psychological mechanism highlights the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal perceptions of success and failure.

>>>>

Meritocracy constitutes a central ideological framework for legitimizing various forms of social inequality [@mijs_unfulfillable_2016; @wilson_role_2003]. Rooted in the belief that rewards and social positions should be allocated based on individual effort and talent, meritocracy operates both as a descriptive account of how societies function and as a normative ideal of how they ought to function. As originally conceptualized by Michael Young [-@young_rise_1962], the term was intended as a critique of a system in which merit-based stratification gives rise to a new form of inequality. However, over time, meritocracy has come to be widely endorsed as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented societies [@mijs_paradox_2021; @sandel_tyranny_2020].

From a sociological perspective, belief in meritocracy extends beyond a cognitive judgment; it constitutes a moral lens through which individuals interpret social and economic disparities [@mijs_paradox_2021]. Individuals who believe that success results from hard work and talent are more likely to regard inequality as legitimate [@tejero-peregrina_perceived_2025; @castillo_percepcion_2019; @batruch_belief_2023]. In contrast, when outcomes are attributed to luck, social origin, or structural barriers, inequality is more often perceived as unjust [@reynolds_perceptions_2014; @hauser_misperceptions_2017]. This distinction becomes especially salient in societies marked by persistent structural inequality, where dominant narratives emphasize personal responsibility while neglecting the role of entrenched disadvantage.

Recent research has highlighted the importance of disaggregating the concept of "meritocratic beliefs," particularly by distinguishing between meritocratic preferences and meritocratic perceptions [@castillo_multidimensional_2023]. While preferences refer to normative ideals about how rewards should be allocated (i.e., based on merit), perceptions refer to subjective evaluations of how meritocracy actually operates in practice (i.e., how rewards are allocated) [@janmaat_subjective_2013]. This distinction allows for a more precise understanding of how individuals make sense of inequality: people may endorse meritocracy normatively while simultaneously recognizing its failure in practice. In stratified societies, the gap between preferences and perceptions is particularly evident, as individuals may continue to uphold the ideal of meritocracy despite growing awareness of unequal opportunities [@mijs_belief_2022].

Within the domain of meritocratic perceptions [@castillo_multidimensional_2023], we further distinguish between two core dimensions: effort-based and talent-based perceptions. This distinction is analytically important, as it captures different mechanisms through which individuals may justify inequality [@young_rise_1962; @reynolds_perceptions_2014]. Effort-based perceptions emphasize hard work and perseverance as the key drivers of success, aligning closely with cultural narratives of individual responsibility and moral deservingness [@dubet_repensar_2011]. In contrast, talent-based perceptions emphasize innate ability and intelligence—traits often regarded as less malleable and more unequally distributed [@young_rise_1962]. We argue that this latter dimension may shape attitudes differently, as talent tends to be seen as a given attribute rather than the result of personal effort.

Meritocratic perceptions function as a normative framework that legitimizes unequal outcomes, particularly when access to social goods is stratified by income or socioeconomic background. Prior studies have consistently shown that individuals who perceive society as meritocratic express lower support for redistribution [@tejero-peregrina_perceived_2025; @hoyt_mindsets_2023], greater legitimization of class inequality [@batruch_belief_2023], increased tolerance for inequality [@day_movin_2017], and lower perceptions of income inequality itself [@castillo_percepcion_2019]. In the regard of market justice, prior studies in Chile have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes in healthcare, education and pensions [@castillo_socialization_2024]. Taken together, and as @busemeyer_positive_2021 argues, meritocratic narratives can operate as feedback mechanisms that shape public opinion and individual well-being by framing welfare outcomes as either deserved or undeserved. This psychological mechanism underscores the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal understandings of success and failure.

### 2025-08-06 resubmission/paper-blinded-resubmit.tex
Up antecedentes meritocracia + render
@juancarloscastillo todo tuyo

  \author{Juan Carlos Castillo}
            \affil{%
                  Departamento de Sociología, Universidad de Chile
              }
          \affil{%
                  Centro de estudios del conflicto y cohesión social
                  (COES)
              }
          \affil{%
                  Núcleo milenio de desigualdades y oportunidades
                  digitales (NUDOS)
              }
        \author{Andreas Laffert}
            \affil{%
                  Instituto de Sociología, Pontificia Universidad
                  Católica de Chile
              }
        \author{Kevin Carrasco}
            \affil{%
                  Centro de estudios del conflicto y cohesión social
                  (COES)
              }
        \author{Julio Iturra-Sanhueza}
            \affil{%
                  International Graduate School of Social Sciencies
                  (BIGSSS), University of Bremen, Germany
              }

\date{}
\begin{document}
\maketitle
\begin{abstract}
Several countries have experienced a shift toward the privatization and
commodification of public goods, welfare policies, and social services.
In Latin America, Chile stands out as a paradigmatic case where this
trend has led to the extensive marketization of essential services. From
a moral economy perspective, the extent to which individuals consider it
fair for access to such services to depend on market criteria has been
conceptualized as market justice preferences. This study investigates
the relationship between perceptions of economic inequality,
meritocratic beliefs, and market justice preferences in Chile between
2016 and 2023. Using six waves of panel data from the Chilean
Longitudinal Social Survey -- ELSOC (\(N_{observations}\) = 8,643;
\(N_{individuals}\) = 1,687), the analysis examines how subjective
assessments of inequality shape attitudes toward the role of merit in
access to key social services such as healthcare, education, and
pensions. The findings show that greater perceived inequality is
associated with lower market justice preferences. However, individuals
who believe that effort is rewarded are more likely to legitimize
existing disparities. In contrast, the perception that talent is
rewarded shows a negative effect on market justice preferences; an
effect that intensifies as perceived inequality increases over time. The
study also considers the influence of major social movements during this
period, which appear to have reshaped public discourse on justice and
fairness. These findings contribute to a deeper understanding of how
distributive beliefs evolve in contexts marked by persistent inequality
and entrenched neoliberal frameworks \newline \textbf{Keywords}:
Inequality perception, meritocracy, commodification, Chile, public
preferences
\end{abstract}

@@

Perceptions of inequality refer to individuals' subjective evaluations
of the extent, causes, and consequences of income and wealth disparities
(\citeproc{ref-akyelken_urban_2020}{Akyelken, 2020}). Unlike objective
measures such as the Gini index, perceived inequality captures how
individuals make sense of distributive hierarchies in their everyday
lives, shaped by reference groups, social comparisons, and information
environments (\citeproc{ref-garcia-castro_perceiving_2020}{García-Castro
et al., 2020}; \citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \&
Treisman, 2018}; \citeproc{ref-mijs_stratified_2016}{Mijs, 2016}).
Scholars have proposed multiple dimensions of perceived inequality,
including its magnitude (how significant are the gaps), vertical
structure (between which groups), the trend over time (increasing or
decreasing), and legitimacy (whether it is just or not)
(\citeproc{ref-engelhardt_what_2018}{Engelhardt \& Wagener, 2018};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These dimensions encompass both cognitive and normative aspects
of perceptions of inequality and can vary across societies and social
groups, depending on exposure, ideology, and personal experience
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}).

@@

Meritocracy constitutes a central ideological framework for legitimizing
different types of social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1962}{1962}), the term was meant to critique a
system in which merit-based stratification becomes a new form of
inequality. However, over time, meritocracy has been widely supported in
many societies as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented economies
(\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}). From a sociological
perspective, the belief in meritocracy is more than a cognitive
assessment; it reflects a moral lens through which individuals interpret
inequality. People who believe that success results from hard work and
talent are more likely to view social and economic disparities as
legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality. Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}).

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they endorse
meritocratic principles. This distinction aligns with recent findings
indicating that individuals distinguish between how merit is perceived
in society and how it should ideally operate, which in turn shapes their
preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}). Meritocratic beliefs serve as symbolic justifications for
unequal outcomes, particularly when access is stratified by income or
social opportunity. Prior studies have shown that individuals who
perceive higher levels of meritocracy tend to express stronger support
for unequal distributions that reflect market outcomes
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

In addition to influencing individual attitudes toward inequality,
meritocratic perceptions can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor (\citeproc{ref-hoyt_mindsets_2023}{Hoyt
et al., 2023}). This reinforces negative stereotypes and reduces empathy
toward individuals from lower socioeconomic backgrounds. Moreover,
Busemeyer et al. (\citeproc{ref-busemeyer_positive_2021}{2021}) argues
that meritocratic narratives can serve as feedback mechanisms that shape
public opinion and well-being by framing individuals' understanding of
welfare outcomes as deserved or undeserved within existing institutional
structures. This psychological mechanism highlights the normative power
of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal perceptions of success and failure.

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-resubmit_files/figure-pdf/fig-alluvial-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-resubmit_files/figure-pdf/fig-meanchange-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-resubmit_files/figure-pdf/fig-interact-1.pdf}

@@

@@

@@

@@

\bibitem[\citeproctext]{ref-mijs_stratified_2016}
Mijs, J. (2016). Stratified {Failure}: {Educational Stratification} and
{Students}' {Attributions} of {Their Mathematics Performance} in 24
{Countries}. \emph{Sociology of Education}, \emph{89}(2), 137--153.
\url{https://doi.org/10.1177/0038040716636434}

@@

@@

>>>>

\date{}
\begin{document}
\maketitle
\begin{abstract}
This study investigates the relationship between perceptions of economic
inequality, meritocratic perceptions, and preferences for market justice
in Chile between 2016 and 2023. Using six waves of panel data from the
Chilean Longitudinal Social Survey - ELSOC (\(N_{observations}\) =
8,643; \(N_{individuals}\) = 1,687), the analysis examines how
subjective assessments of inequality shape attitudes toward the role of
merit in access to key social services such as healthcare, education,
and pensions. Results indicate that rising perceptions of inequality are
associated with lower support for market justice; however, individuals
with strong meritocratic convictions are more likely to legitimize
existing disparities. The study also considers the influence of major
social movements during this period, which appear to have reshaped
public discourse and perceptions of fairness. These findings contribute
to a deeper understanding of how beliefs about justice and equity evolve
in contexts marked by persistent inequality and entrenched
market-oriented frameworks. \newline \textbf{Keywords}: Economic
inequality, meritocracy, market justice, Chile, public preferences,
inequality perception
\end{abstract}

@@

Perceptions of inequality refer to individuals' subjective evaluations
of the extent, causes, and consequences of income and wealth disparities
(\citeproc{ref-akyelken_urban_2020}{Akyelken, 2020}). Unlike objective
measures such as the Gini index, perceived inequality captures how
individuals make sense of distributive hierarchies in their everyday
lives, shaped by reference groups, social comparisons, and information
environments (\citeproc{ref-garcia-castro_perceiving_2020}{García-Castro
et al., 2020}; \citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \&
Treisman, 2018}; \citeproc{ref-mijs_stratified_2016}{Mijs, 2016a}).
Scholars have proposed multiple dimensions of perceived inequality,
including its magnitude (how significant are the gaps), vertical
structure (between which groups), the trend over time (increasing or
decreasing), and legitimacy (whether it is just or not)
(\citeproc{ref-engelhardt_what_2018}{Engelhardt \& Wagener, 2018};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These dimensions encompass both cognitive and normative aspects
of perceptions of inequality and can vary across societies and social
groups, depending on exposure, ideology, and personal experience
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}).

@@

Meritocracy constitutes a central ideological framework for legitimizing
various forms of social inequality
(\citeproc{ref-mijs_unfulfillable_2016}{Mijs, 2016b};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). Rooted in the belief
that rewards and social positions should be allocated based on
individual effort and talent, meritocracy operates both as a descriptive
account of how societies function and as a normative ideal of how they
ought to function. As originally conceptualized by Michael Young
(\citeproc{ref-young_rise_1962}{1962}), the term was intended as a
critique of a system in which merit-based stratification gives rise to a
new form of inequality. However, over time, meritocracy has come to be
widely endorsed as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented societies
(\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}).

From a sociological perspective, belief in meritocracy extends beyond a
cognitive judgment; it constitutes a moral lens through which
individuals interpret social and economic disparities
(\citeproc{ref-mijs_paradox_2021}{Mijs, 2021}). Individuals who believe
that success results from hard work and talent are more likely to regard
inequality as legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et
al., 2023}; \citeproc{ref-castillo_percepcion_2019}{Castillo et al.,
2012}; \citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina
et al., 2025}). In contrast, when outcomes are attributed to luck,
social origin, or structural barriers, inequality is more often
perceived as unjust (\citeproc{ref-hauser_misperceptions_2017}{Hauser \&
Norton, 2017}; \citeproc{ref-reynolds_perceptions_2014}{Reynolds \&
Xian, 2014}). This distinction becomes especially salient in societies
marked by persistent structural inequality, where dominant narratives
emphasize personal responsibility while neglecting the role of
entrenched disadvantage.

Recent research has highlighted the importance of disaggregating the
concept of ``meritocratic beliefs,'' particularly by distinguishing
between meritocratic preferences and meritocratic perceptions
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
While preferences refer to normative ideals about how rewards should be
allocated (i.e., based on merit), perceptions refer to subjective
evaluations of how meritocracy actually operates in practice (i.e., how
rewards are allocated) (\citeproc{ref-janmaat_subjective_2013}{Janmaat,
2013}). This distinction allows for a more precise understanding of how
individuals make sense of inequality: people may endorse meritocracy
normatively while simultaneously recognizing its failure in practice. In
stratified societies, the gap between preferences and perceptions is
particularly evident, as individuals may continue to uphold the ideal of
meritocracy despite growing awareness of unequal opportunities
(\citeproc{ref-mijs_belief_2022}{Mijs et al., 2022}).

Within the domain of meritocratic perceptions
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}),
we further distinguish between two core dimensions: effort-based and
talent-based perceptions. This distinction is analytically important, as
it captures different mechanisms through which individuals may justify
inequality (\citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian,
2014}; \citeproc{ref-young_rise_1962}{Young, 1962}). Effort-based
perceptions emphasize hard work and perseverance as the key drivers of
success, aligning closely with cultural narratives of individual
responsibility and moral deservingness
(\citeproc{ref-dubet_repensar_2011}{Dubet, 2011}). In contrast,
talent-based perceptions emphasize innate ability and
intelligence---traits often regarded as less malleable and more
unequally distributed (\citeproc{ref-young_rise_1962}{Young, 1962}). We
argue that this latter dimension may shape attitudes differently, as
talent tends to be seen as a given attribute rather than the result of
personal effort.

Meritocratic perceptions function as a normative framework that
legitimizes unequal outcomes, particularly when access to social goods
is stratified by income or socioeconomic background. Prior studies have
consistently shown that individuals who perceive society as meritocratic
express lower support for redistribution
(\citeproc{ref-hoyt_mindsets_2023}{Hoyt et al., 2023};
\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}), greater legitimization of class inequality
(\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023}), increased
tolerance for inequality (\citeproc{ref-day_movin_2017}{Day \& Fiske,
2017}), and lower perceptions of income inequality itself
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012}). In the
regard of market justice, prior studies in Chile have shown that
individuals who perceive higher levels of meritocracy tend to express
stronger support for unequal distributions that reflect market outcomes
in healthcare, education and pensions
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).
Taken together, and as Busemeyer et al.
(\citeproc{ref-busemeyer_positive_2021}{2021}) argues, meritocratic
narratives can operate as feedback mechanisms that shape public opinion
and individual well-being by framing welfare outcomes as either deserved
or undeserved. This psychological mechanism underscores the normative
power of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal understandings of success and failure.

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-blinded-resubmit_files/figure-pdf/fig-alluvial-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-blinded-resubmit_files/figure-pdf/fig-meanchange-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper-blinded-resubmit_files/figure-pdf/fig-interact-1.pdf}

@@

\bibitem[\citeproctext]{ref-day_movin_2017}
Day, M. V., \& Fiske, S. T. (2017). Movin' on {Up}? {How Perceptions} of
{Social Mobility Affect Our Willingness} to {Defend} the {System}.
\emph{Social Psychological and Personality Science}, \emph{8}(3),
267--274. \url{https://doi.org/10.1177/1948550616678454}

@@

\bibitem[\citeproctext]{ref-dubet_repensar_2011}
Dubet, F. (2011). \emph{{Repensar la justicia social}} (Sexta
Edici{ó}n). Siglo XXI.

@@

\bibitem[\citeproctext]{ref-hauser_misperceptions_2017}
Hauser, O. P., \& Norton, M. I. (2017). ({Mis})perceptions of
inequality. \emph{Current Opinion in Psychology}, \emph{18}, 21--25.
\url{https://doi.org/10.1016/j.copsyc.2017.07.024}

@@

\bibitem[\citeproctext]{ref-mijs_stratified_2016}
Mijs, J. (2016a). Stratified {Failure}: {Educational Stratification} and
{Students}' {Attributions} of {Their Mathematics Performance} in 24
{Countries}. \emph{Sociology of Education}, \emph{89}(2), 137--153.
\url{https://doi.org/10.1177/0038040716636434}

\bibitem[\citeproctext]{ref-mijs_unfulfillable_2016}
Mijs, J. (2016b). The {Unfulfillable Promise} of {Meritocracy}: {Three
Lessons} and {Their Implications} for {Justice} in {Education}.
\emph{Social Justice Research}, \emph{29}(1), 14--34.
\url{https://doi.org/10.1007/s11211-014-0228-0}

@@

\bibitem[\citeproctext]{ref-mijs_belief_2022}
Mijs, J., Daenekindt, S., de Koster, W., \& van der Waal, J. (2022).
Belief in {Meritocracy Reexamined}: {Scrutinizing} the {Role} of
{Subjective Social Mobility}. \emph{Social Psychology Quarterly},
\emph{85}(2), 131--141. \url{https://doi.org/10.1177/01902725211063818}

@@

\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
Reynolds, J., \& Xian, H. (2014). Perceptions of meritocracy in the land
of opportunity. \emph{Research in Social Stratification and Mobility},
\emph{36}, 121--137. \url{https://doi.org/10.1016/j.rssm.2014.03.001}
